---
title: "HW06"
author: "Elliot Smith, Eugen Hruska, Varun Suriyanaranana"
date: "3/9/2018"
output: pdf_document
---

```{r, echo = FALSE}

##### Workspace Preparation

## Load in packages

library(ggplot2)
library(knitr)
library(png)
library(grid)

## Load in euclidean distance function

euclidean_distance <- function(p, q){
    sqrt(sum((p - q)^2))
}

```

# Problem 2.1

```{r, echo = FALSE}

##### Problem 2.1

## Load in the iris training data

train_data <- read.table("~/Documents/Rice_University/Spring_2018/NML502/HW04_Part2/iris-train.txt", skip = 8)
train_attr <- train_data[c(TRUE, FALSE), ]
train_attr$V1 <- as.numeric(as.character(train_attr$V1))
train_cats <- train_data[c(FALSE, TRUE), ][ ,2:4]

data <- cbind(train_attr, train_cats)
names(data) <- c("Sep_L", "Sep_W", "Pet_L", "Pet_W", "Setosa", "Versacolor", "Virginica")

## Scale the data

scaled_data <- apply(data[, 1:4], 2, function(x) {(x - mean(x)) / sd(x)})

## Perform the PCA

iris_pca <- prcomp(scaled_data)

eigen_vals <- iris_pca$sdev
eigen_vecs <- iris_pca$rotation

## Confirm we get the identity matrix

id_conf <- iris_pca$rotation %*% t(iris_pca$rotation)

```

## Eigenvectors

```{r, echo = FALSE}

kable(eigen_vecs)

```

## Eigenvalues

```{r, echo = FALSE}

kable(matrix(eigen_vals, nrow = 1))

```

## Eigenvector Matrix x Eigenvector Matrix Transposed

```{r, echo = FALSE}

kable(id_conf)

```

# Problem 2.2

```{r, echo = FALSE}

learn_gha <- function(x, n, num_outputs, num_layers, ler_rate, weights) {
    
    learn_matrices <- list()
    learn_step <- numeric()
    euc_dist <- numeric()
    
    for (i in 1:n) {
        
        rand_ind <- sample(1:75, 1)
        
        y <- weights %*% matrix(x[rand_ind, ], ncol = 1)
        
        temp_mat <- matrix(0, nrow = 4, ncol = 4)
        
        for (j in 1:length(y)) {
            
            if (j == 1) {
                
                temp_mat[j, ] <- (y[j] * y[j] * weights[j, ])
                
            } else if (j == 2) {
                
                temp_mat[j, ] <- (y[j] * y[j - 1] * weights[j - 1, ]) + (y[j] * y[j] * weights[j, ])
                
            } else if (j == 3) {
                
                temp_mat[j, ] <- (y[j] * y[j - 2] * weights[j - 2, ]) + (y[j] * y[j - 1] * weights[j - 1, ]) + (y[j] * y[j] * weights[j, ])
                
            } else if (j == 4) {
                
                temp_mat[j, ] <- (y[j] * y[j - 3] * weights[j - 3, ]) + (y[j] * y[j - 2] * weights[j - 2, ]) + (y[j] * y[j - 1] * weights[j - 1, ]) + (y[j] * y[j] * weights[j, ])
                
            }
            
        }
        
        weights <- weights +
            ler_rate * (matrix(y, ncol = 1) %*% matrix(x[rand_ind, ], nrow = 1)) -
            ler_rate * temp_mat
        
        if (i %% 100 == 0) {
            
            learn_step[length(learn_step) + 1] <- i
            learn_matrices[[length(learn_matrices) + 1]] <- weights %*% t(weights)
            euc_dist[length(euc_dist) + 1] <- euclidean_distance(id_matrix, weights %*% t(weights))
            
        }
        
    }
    
    ggplot() +
        geom_line(aes(x = learn_step, y = euc_dist)) +
        labs(x = "Learning Step", y = "Euclidean Distance", title = "Matrix Distance",
             subtitle = "Distance between Weight Matrix times its Transpose and the Identity Matrix")
    
    final_weights <<- weights
    
}

```

## Learning History

```{r, echo = FALSE}

## Set the network parameters

x <- scaled_data
n <- 1500000
num_outputs <- c(4, 4)
num_layers <- 1
ler_rate <- 0.0005

## Build the weight matrix

weights <- list()

weights <- matrix(runif(num_outputs[num_layers] * num_outputs[num_layers + 1], min = -0.1, max = 0.1),
                  nrow = num_outputs[num_layers + 1],
                  ncol = num_outputs[num_layers])

init_weights <- weights

## Set the identity matrix

id_matrix <- diag(x = 4)

## Learn the network

learn_gha(x, n, num_outputs, num_layers, ler_rate, weights)

## Read in the plot

img <- readPNG("~/Documents/Rice_University/Spring_2018/NML502/HW06/Plot1.png")
grid.raster(img)

```

# Comparison 2.1 to 2.2

## Initial Weight Matrix

```{r, echo = FALSE}

kable(init_weights)

```

## Learned Weight Matrix

```{r, echo = FALSE}

kable(t(final_weights))

```

## Network Information

* Number of Learning Steps: 1500000
* Learning Rate: 0.0005
* Learning History: See graph above (defined by Euclidean distance between weight matrix times its transpose and the Identity matrix)
* Preprocessing: All data is centered (zero mean)
* Error Measure: Euclidean distance

# Code Appendix

```{r, eval = FALSE}

##### Workspace Preparation

## Load in packages

library(ggplot2)

## Load in euclidean distance function

euclidean_distance <- function(p, q){
    sqrt(sum((p - q)^2))
}

##### Problem 2.1

## Load in the iris training data

train_data <- read.table("~/Documents/Rice_University/Spring_2018/NML502/HW04_Part2/iris-train.txt",
                         skip = 8)
train_attr <- train_data[c(TRUE, FALSE), ]
train_attr$V1 <- as.numeric(as.character(train_attr$V1))
train_cats <- train_data[c(FALSE, TRUE), ][ ,2:4]

data <- cbind(train_attr, train_cats)
names(data) <- c("Sep_L", "Sep_W", "Pet_L", "Pet_W", "Setosa", "Versacolor", "Virginica")

## Scale the data

scaled_data <- apply(data[, 1:4], 2, function(x) {(x - mean(x)) / sd(x)})

## Perform the PCA

iris_pca <- prcomp(scaled_data)

eigen_vals <- iris_pca$sdev
eigen_vecs <- iris_pca$rotation

## Confirm we get the identity matrix

iris_pca$rotation %*% t(iris_pca$rotation)



##### Problem 2.2

## Construct the learning function

learn_gha <- function(x, n, num_outputs, num_layers, ler_rate, weights) {
    
    learn_matrices <- list()
    learn_step <- numeric()
    euc_dist <- numeric()
    
    for (i in 1:n) {
        
        rand_ind <- sample(1:75, 1)
        
        y <- weights %*% matrix(x[rand_ind, ], ncol = 1)
        
        temp_mat <- matrix(0, nrow = 4, ncol = 4)
        
        for (j in 1:length(y)) {
            
            if (j == 1) {
                
                temp_mat[j, ] <- (y[j] * y[j] * weights[j, ])
                
            } else if (j == 2) {
                
                temp_mat[j, ] <- (y[j] * y[j - 1] * weights[j - 1, ]) +
                    (y[j] * y[j] * weights[j, ])
                
            } else if (j == 3) {
                
                temp_mat[j, ] <- (y[j] * y[j - 2] * weights[j - 2, ]) +
                    (y[j] * y[j - 1] * weights[j - 1, ]) +
                    (y[j] * y[j] * weights[j, ])
                
            } else if (j == 4) {
                
                temp_mat[j, ] <- (y[j] * y[j - 3] * weights[j - 3, ]) +
                    (y[j] * y[j - 2] * weights[j - 2, ]) +
                    (y[j] * y[j - 1] * weights[j - 1, ]) +
                    (y[j] * y[j] * weights[j, ])
                
            }
            
        }
        
        weights <- weights +
            ler_rate * (matrix(y, ncol = 1) %*% matrix(x[rand_ind, ], nrow = 1)) -
            ler_rate * temp_mat
        
        if (i %% 100 == 0) {
            
            learn_step[length(learn_step) + 1] <- i
            learn_matrices[[length(learn_matrices) + 1]] <- weights %*% t(weights)
            euc_dist[length(euc_dist) + 1] <- euclidean_distance(id_matrix, weights %*% t(weights))
            
        }
        
    }
    
    ggplot() +
        geom_line(aes(x = learn_step, y = euc_dist)) +
        labs(x = "Learning Step", y = "Euclidean Distance", title = "Matrix Distance",
             subtitle = "Distance between Weight Matrix times its Transpose and the Identity Matrix")
    
}

## Set the network parameters

x <- scaled_data
n <- 1500000
num_outputs <- c(4, 4)
num_layers <- 1
ler_rate <- 0.0005

## Build the weight matrix

weights <- list()

weights <- matrix(runif(num_outputs[num_layers] * num_outputs[num_layers + 1], min = -0.1, max = 0.1),
                  nrow = num_outputs[num_layers + 1],
                  ncol = num_outputs[num_layers])

## Set the identity matrix

id_matrix <- diag(x = 4)

## Learn the network

learn_gha(x, n, num_outputs, num_layers, ler_rate, weights)

```















